Optimal design of experiment

### 2024-01-21 SFE_article_Design_of_Experiment/Sections/Optimal_design_of_experiment.tex
21.01

\documentclass[]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption}

%opening
\title{Optimal design of experiment}
\author{Oliwer Sliczniuk}

\begin{document}

\maketitle

\begin{abstract}
Notes about the optimal design of an experiment
\end{abstract}

\section{What information do we have?}

\begin{itemize}
	\item Process model: $\dot{x}=f(t,x,p)$
	\item Measurement equation: $\dot{y}=g(x)$
	\item Likelihood function with respect to the dataset
	\item Parameter estimation
	\item Sensitivity equations: $\dot{S} = J_x \dot{x} + J_p$
\end{itemize}

\section{Maximum likelihood}
Let $y^s$ be the vector of all experimental results (random vector) to be used to estimate parameters $p$ and model output $y^m(p)$. The vector of the corresponding quantities computed by the model $\dot{x}=f(t,x,p)$. For a parallel model, the parameters of which are to be estimated from measurements of an output vector $\dot{y}=g(x)$. We shall call any optimizer of $j$, i.e., any $\hat{p}$ that corresponds to an optimal value of the cost function $j$, an estimate of $p$ in the sense of $j$.

@@

\begin{equation}
	j_{ml} = \pi_y (y^s|p) 
\end{equation}

If $p$ were fixed, $\pi_y(y^s|p)$ would be the probability density of the random vector $y^s$ generated by a model with parameters $p$. Here, to the contrary, $y^s$ is fixed and corresponds to the observations. Considered as a function of $p$, $\pi_y(y^s|p)$ is then called the likelihood of $y^s$. The maximum-likelihood method looks for the parameter vector $p$ value that gives the observed data the highest likelihood. In practice, it is often easier to look for $\hat{p}_{ml}$ by maximizing the log-likelihood function, yielding the same estimate since the logarithm function is monotonically increasing.

\begin{equation}
	j_{ml} = \ln ( \pi_y (y^s|p) )
\end{equation}

@@

\begin{equation}
	y(t_i) = y^m(t_i, p^*) + \epsilon_i, \qquad i=1,...,n_t
\end{equation}

where the vector $y^m(t_i. p^*)$ is the output of a deterministic model, $p^*$ is the true value of the parameter vector, and $\epsilon_i$ belongs to a sequence of independent random variables with probability density $\pi_{\epsilon}(\epsilon_i)$. Since the $\epsilon_i$ are independent

\begin{equation}
	\pi_{\epsilon}(\epsilon_1, \epsilon_2, ..., \epsilon_{n_t}) = \prod_{i=1}^{n_t} \pi_{\epsilon_i}(\epsilon_i)
\end{equation}

@@

\begin{equation}
	e^y(t_i, p) = y(t_i) - y^m(t_i, p)
\end{equation}

For the true value of the parameters, it satisfies $e^y(t_i, p^*)=\epsilon_i$. and, since $y^m$ is deterministic, $\pi_{y_i} (y^s(t_i)|p) = \pi_{\epsilon_i} (e^y(t_i)|p)$. The likelihood of the $n_t$ observations can be written as

\begin{equation}
	\pi_y (y^s|p) = \prod_{i=1}^{n_t} \pi_{\epsilon_i}(e^y(t_i,p)) =  \prod_{i=1}^{n_t} \pi_{\epsilon_i}(y(t_i) - y^m(t_i, p))
\end{equation}

@@

\begin{equation}
	\pi_y (y^s|p) = \prod_{i=1}^{n_t} \frac{1}{ \sqrt{2\pi\sigma_{t_i}^2} } \exp \left( -\frac{1}{2} \left( \frac{y(t_i) - y^m(t_i, p)}{\sigma_{t_i}} \right)^2 \right)
\end{equation}

@@

\begin{equation}
	\ln (\pi_y (y^s|p)) = (\text{term independent of p}) - \frac{1}{2} \sum_{i=1}^{n_t}  \left( \frac{y(t_i) - y^m(t_i, p)}{\sigma_{t_i}} \right)^2
\end{equation}

@@

\begin{equation}
	\frac{\partial}{\partial p} \ln (\pi_y (y^s|p)) =  \sum_{i=1}^{n_t}  \left[ \left( \frac{y(t_i) - y^m(t_i, p)}{\sigma_{t_i}^2} \right) \frac{\partial y^m(t_i, p)}{\partial p} \right]
\end{equation}

\section{Fisher information}

@@

\begin{align}
	F(p) &= - \mathop{\mathbb{E}}_{y^s|p} \left[ \frac{\partial^2 \ln (\pi_y (y^s|p))}{\partial p \partial p^T} \right] = \mathop{\mathbb{E}}_{y^s|p} \left[ \frac{\partial \ln (\pi_y (y^s|p))}{\partial p} \frac{\partial \ln (\pi_y (y^s|p))}{\partial p^T} \right] \nonumber \\
	&= \mathop{\mathbb{E}}_{y^s|p} \left[ \sum_{k=1}^{n_t} \left( \frac{y(t_k) - y^m(t_k, p)}{\sigma_{t_k}^2} \frac{\partial y^m(t_k, p)}{\partial p} \right) \times  \sum_{i=1}^{n_t} \left( \frac{y(t_i) - y^m(t_i, p)}{\sigma_{t_i}^2} \frac{\partial y^m(t_i, p)}{\partial p^T} \right) \right] \nonumber \\
	&= \mathop{\mathbb{E}}_{y^s|p} \left[ \sum_{k=1}^{n_t} \sum_{i=1}^{n_t} \left[ \left( \frac{y(t_k) - y^m(t_k, p)}{\sigma_{t_k}^2} \frac{\partial y^m(t_k, p)}{\partial p} \right) \times \left( \frac{y(t_i) - y^m(t_i, p)}{\sigma_{t_i}^2} \frac{\partial y^m(t_i, p)}{\partial p^T} \right) \right] \right] \nonumber \\
	&= \sum_{k=1}^{n_t} \sum_{i=1}^{n_t} \mathop{\mathbb{E}}_{y^s|p} \left[ \left( \frac{y(t_k) - y^m(t_k, p)}{\sigma_{t_k}^2} \frac{\partial y^m(t_k, p)}{\partial p} \right) \times \left( \frac{y(t_i) - y^m(t_i, p)}{\sigma_{t_i}^2} \frac{\partial y^m(t_i, p)}{\partial p^T} \right) \right] \nonumber \\
	&= \sum_{k=1}^{n_t} \sum_{i=1}^{n_t} \frac{1}{\sigma_{t_k}^2} \frac{\partial y^m(t_k, p)}{\partial p} \frac{1}{\sigma_{t_i}^2} \frac{\partial y^m(t_i, p)}{\partial p^T} \mathop{\mathbb{E}}_{y^s|p} \left[ \left( y(t_k) - y^m(t_k, p) \right) \times \left( y(t_i) - y^m(t_i, p) \right) \right] 
\end{align}

Considering the mismatch between the dataset and the model output, the expectation $\mathop{\mathbb{E}}_{y^s|p}$ can be redefined as $\mathop{\mathbb{E}}_{y^s|p} \left[ \left( y(t_i) - y^m(t_i, p) \right) \left( y(t_k) - y^m(t_k, p) \right) \right] = \mathop{\mathbb{E}}_{y^s|p} \left[ \epsilon_i \epsilon_k \right]$. For $i=k$, the expectation of $\epsilon_i$ is the variance of the noise $\sigma_{t_i}^2$. Analogously, the expectation becomes 0 if $i \neq k$ because the measurement noises are independent at different times.

\begin{equation}
	\mathop{\mathbb{E}}_{y^s|p} \left[ \left( y(t_i) - y^m(t_k, p) \right) \left( y(t_i) - y^m(t_i, p) \right) \right] = \sigma_{t_i}^2 \delta_{ik}
\end{equation}

@@

\begin{equation}
	F(p) = \sum_{i=1}^{n_t} \left( \frac{1}{\sigma_{t_i}^2} \frac{\partial y^m(t_i, p)}{\partial p} \frac{\partial y^m(t_i, p)}{\partial p^T} \right)
\end{equation}

@@

\begin{equation}
	F(p) = \frac{\partial y^m(t, p)}{\partial p} \begin{bmatrix}
		\frac{1}{\sigma_{t_1}^2} & 0 & 0\\
		0 & \ddots & 0 \\
		0 & 0 & \frac{1}{\sigma_{t_{n_t}}^2} 
	\end{bmatrix}\ \frac{\partial y^m(t, p)}{\partial p^T} 
\end{equation}

\section{Cramer-Rao inequality}

@@

\begin{equation}
	P = \mathop{\mathbb{E}}_{y^s|p^*} \left[ \left( \hat{p}(y^s) - p^* \right) \left( \hat{p}(y^s) - p^* \right)^T \right]
\end{equation}

which quantifies how the estimates are spread around the true value $p^*$. One would like the estimates to be as concentrated as possible around this true value. An estimator $\hat{p}_1$ with covariance matrix $P_1$ is said to be more efficient than an estimator $\hat{p}_2$ with covariance matrix $P_2$ if $P_1 < P_2$, that is if $P_2 - P_1$ is positive-definite (i.e.if all the eigenvalues of $P_2-P_1$ arc strictly positive). Since estimators with high efficiency are desirable, a natural request is to make P as small as possible. The Cramer-Rao inequality provides a lower bound to what can be achieved.

@@

\begin{itemize}
	\item the set of all data vectors $y^s$ with $\pi_y(y^s|p) > 0$ does not depend on $p$
	\item $\frac{\partial \pi_y(y^s|p)}{\partial p_i}~\left(i=1,2,...,n_p\right)$ is absolutely integrable
	\item $\mathop{\mathbb{E}}_{y^s|p} \left[ \frac{\partial \ln (\pi_y (y^s|p))}{\partial p} \frac{\partial \ln (\pi_y (y^s|p))}{\partial p^T} \right]$ exists and is invertible
\end{itemize}

@@

\begin{equation}
	P \geq F^{-1}(p^*)
\end{equation}

@@

\section{Optimal experimental design}

@@

\begin{equation}
	\Xi^* = \arg~\underset{\Xi \in \bar{\Xi}}{\text{opt}}~j\left(\Xi\right)
\end{equation}

@@

\begin{equation}
	j(\Xi) = \phi\left[ F(p, \Xi) \right]
\end{equation}

@@

\begin{align}
	\phi_k(F) &= \left[\frac{1}{n_p} \text{trace}\left( QF^{-1}Q^T \right)^k \right]^{1/k} \quad &if \det F \neq 0 \nonumber \\
	\phi_k(F) &= \infty \quad &if \det F = 0
\end{align}

@@

\begin{equation}
	j_L(\Xi) = \text{trace} \left[ Q^TQF^{-1}(p,\Xi) \right]
\end{equation}

and choosing $Q = \textbf{I}_{n_p}$ then corresponds to the A-optimality cost function. An A-optimal experiment minimizes the sum of the squares of the lengths or the axes of asymptotic confidence ellipsoids. Choosing $Q$ diagonal, with $[Q]_{ii} = 1/p_i$ corresponds to C-optimality, which is connected with the relative precision of estimates. Taking Q to be a row vector leads to c-optimality. Taking $Q = \textbf{I}_{n_p}$ and $k = \infty$ corresponds to E-optimality; E-optimal design maximizes the smallest eigenvalues of the Fisher information matrix and thus minimizes the length of the largest axis of the asymptotic confidence ellipsoids. The most widely used optimality criterion has $k = 0$, $Q = \textbf{I}_{n_p}$ , requiring minimization of $\det F^{-1}(p, \Xi)$, or, equivalently, maximization of

\begin{equation}
	j_D(\Xi) = \det F(p, \Xi)
\end{equation}

@@

\section{Problem formulation}

@@

\begin{equation}
	\begin{aligned} 
		&\Xi^* &= \arg \min_{ T^0, T^{in} \in \Xi} -j_D(\Xi,y)  \\
		&\text{subject to}
		& \dot{x} = f(x,t,\theta;\Xi) \\
		&& \dot{y} = g(x(t)) \\
		&& 40^\circ C \leq T^{in}(t) \leq 50^\circ C \\
		&& T^{0} = T^{in}(t=0) \\
	\end{aligned}
\end{equation}

The non-linear optimization problem is solved by applying single-shooting to obtain the trajectory under assumption of piece-wise constant controls. The CasADi framework is used as the interface for the IPOPT solver. Each simulation problem is solved multiple time starting from different initial points and the solution with the best score is consider as the optimal solution.

\section{Results and discussion}
The D-optimal Design of the Experiment is used to define the way to run the experiment to validate the hypothesis presented in $({\color{red}article~1})$, which are the correlations for $D_i$ and $\Gamma$. It is assumed that the experiment is performed on the same equipment as in the previous work, so all the properties of the machine and the biomass stay the same. The flow rate and the pressure are considered to be constant during the whole experiment. In the previous work, the pressure range was from 200 bar to 300 bar; in this experiment, the pressure is assumed to be 250 bar. The initial and the inlet temperatures are found by solving the optimization problem. The optimization problem was solved 69 times, starting from random initial conditions between $40^\circ C$ and $50^\circ C$. Figure \ref{fig:monet_carlo_P250} summarize solutions to all the optimization problem. The x-axis corresponds to the initial value of the D-optimal score, while the y-axis represents the D-optimal score's final value. As the optimization problem is formulated as the minimization, the lowest negative value is considered the best.

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.80\textwidth, trim = 0cm 0cm 0cm 0cm, clip]{Multiple_shot_DOE_P250.pdf}
	\caption{Comparison between initial and optimal solutions}
	\label{fig:monet_carlo_P250}
\end{figure}

Based on Figure \ref{fig:monet_carlo_P250}, five best results were selected and the the corresponding inlet temperatures are presented on Figure \ref{fig:Best_T_in_P250}. Corresponding outlet temperatures are presented in Figure \ref{fig:Best_T_out_P250}. The inlet and outlet temperature profiles difference from each other due to thermal diffusion. For all five cases, the inlet profiles overlap with each other.

\begin{figure}[!h]
	\centering
	\begin{subfigure}[h]{0.5\textwidth}
		\centering
		\includegraphics[width=\textwidth,trim = 0cm 0cm 0cm 0cm, clip]{Best_T_in_P250.pdf}
		\caption{Inlet temperature profile}
		\label{fig:Best_T_in_P250}
	\end{subfigure}%
	~ 
	\begin{subfigure}[h]{0.5\textwidth}
		\centering
		\includegraphics[width=\textwidth,trim = 0cm 0cm 0cm 0cm, clip]{Best_T_out_P250.pdf}
		\caption{Outlet temperature profile}
		\label{fig:Best_T_out_P250}
	\end{subfigure}
	\caption{Temperature profiles}
	\label{fig:DOE_results}
\end{figure}

It can be observed that all the results are very similar to each other and represent the same pattern. The optimizer found that the best score, given the assumptions explained above, can be obtained by the square waves. This pattern is characterized by periods where the temperature remains constant at a particular value, followed by sudden and steep changes to a new constant value. In this graph, the temperature alternates between roughly $313$ K and $323$ K.

At the beginning of the experiment, the temperature rapidly rises to $323$ K and stays flat at this high level for about 20 minutes, indicating that the supercritical $CO_2$ has reached a stable temperature. Following this period, there is a sharp decline to a lower stable temperature of around $313$ K, where it again remains constant for a similar duration. After maintaining the lower temperature, the pattern repeats itself with another sharp increase back to $323$ K, followed by a period of stability and then another drop.

\end{document}

>>>>

\documentclass[../Article_Design_of_Experiment.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

	The design of experiments (DOE) also known as experiment design or experimental design, is the design of any task that aims to describe and explain the variation of information under conditions that are hypothesized to reflect the variation. The term is generally associated with experiments in which the design introduces conditions that directly affect the variation, but may also refer to the design of quasi-experiments, in which natural conditions that influence the variation are selected for observation.

	\subsubsection{Maximum likelihood}
	%Let $Y$ be the vector of all experimental results (random vector) to be used to estimate parameters $p$ (where $p$ is the subset of parameter space $\Theta$) and model output $y(p)$. The vector of the corresponding quantities computed by the model $\dot{x}(t)=G(x(t);p)$. For a parallel model, the parameters of which are to be estimated from measurements of an output vector $\dot{y}=g(x)$. We shall call any optimizer of $j$, i.e., any $\hat{p}$ that corresponds to an optimal value of the cost function $j$, an estimate of $p$ in the sense of $j$.

@@

	{\footnotesize
	\begin{equation}
		j_{ml}(p) = \pi_y (Y|p) 
	\end{equation} }

	If $p$ were fixed, $\pi_y(Y|p)$ would be the probability density of the random vector $Y$ generated by a model with parameters $p$. Here, to the contrary, $Y$ is fixed and corresponds to the observations. Considered as a function of $p$, $\pi_y(Y|p)$ is then called the likelihood of $Y$. The maximum-likelihood method looks for the parameter vector $p$ value that gives the observed data the highest likelihood. In practice, it is often easier to look for $\hat{p}_{ml}$ by maximizing the log-likelihood function, yielding the same estimate since the logarithm function is monotonically increasing.

	{\footnotesize
	\begin{equation}
		j_{ml}(p) = \ln ( \pi_y (Y|p) )
	\end{equation} }

@@

	{\footnotesize
	\begin{equation}
		Y(t_i) = y(t_i, p^*) + \epsilon_i, \qquad i=1,...,n_t
	\end{equation} }

	where the vector $y(t_i, p^*)$ is the output of a deterministic model, $p^*$ is the true value of the parameter vector, and $\epsilon_i$ belongs to a sequence of independent random variables with probability density $\pi_{\epsilon}(\epsilon_i)$. Since the $\epsilon_i$ are independent

	{\footnotesize
	\begin{equation}
		\pi_{\epsilon}(\epsilon_1, \epsilon_2, ..., \epsilon_{n_t}) = \prod_{i=1}^{n_t} \pi_{\epsilon_i}(\epsilon_i)
	\end{equation} }

@@

	{\footnotesize
	\begin{equation}
		e^y(t_i, p) = Y(t_i) - y(t_i, p)
	\end{equation} }

	For the true value of the parameters, it satisfies $e^y(t_i, p^*)=\epsilon_i$. and, since $y$ is deterministic, $\pi_{y_i} (Y(t_i)|p) = \pi_{\epsilon_i} (e^Y(t_i)|p)$. The likelihood of the $n_t$ observations can be written as

	{\footnotesize
	\begin{equation}
		\pi_y (Y|p) = \prod_{i=1}^{n_t} \pi_{\epsilon_i}(e^y(t_i,p)) =  \prod_{i=1}^{n_t} \pi_{\epsilon_i}(Y(t_i) - y(t_i, p))
	\end{equation} }

@@

	{\footnotesize
	\begin{equation}
		\pi_y (Y|p) = \prod_{i=1}^{n_t} \frac{1}{ \sqrt{2\pi\sigma_{t_i}^2} } \exp \left( -\frac{1}{2} \left( \frac{Y(t_i) - y(t_i, p)}{\sigma_{t_i}} \right)^2 \right)
	\end{equation} }

@@

	{\footnotesize
	\begin{equation}
		\ln (\pi_y (Y|p)) = (\text{term independent of p}) - \frac{1}{2} \sum_{i=1}^{n_t}  \left( \frac{Y(t_i) - y(t_i, p)}{\sigma_{t_i}} \right)^2
	\end{equation} }

@@

	{\footnotesize
	\begin{equation}
		\frac{\partial}{\partial p} \ln (\pi_y (Y|p)) =  \sum_{i=1}^{n_t}  \left[ \left( \frac{Y(t_i) - y(t_i, p)}{\sigma_{t_i}^2} \right) \frac{\partial y(t_i, p)}{\partial p} \right]
	\end{equation} }

	\subsubsection{Fisher information}

@@

	{\scriptsize
	\begin{equation}
		\begin{aligned}
		&\text{FIM(p)} = - \mathop{\mathbb{E}}_{Y|p} \left[ \frac{\partial^2 \ln (\pi_y (Y|p))}{\partial p \partial p^\top} \right] = \mathop{\mathbb{E}}_{Y|p} \left[ \frac{\partial \ln (\pi_y (Y|p))}{\partial p} \frac{\partial \ln (\pi_y (Y|p))}{\partial p^\top} \right] \nonumber \\
		&= \mathop{\mathbb{E}}_{Y|p} \left[ \sum_{k=1}^{n_t} \left( \frac{Y(t_k) - y(t_k, p)}{\sigma_{t_k}^2} \frac{\partial y(t_k, p)}{\partial p} \right) \times  \sum_{i=1}^{n_t} \left( \frac{Y(t_i) - y(t_i, p)}{\sigma_{t_i}^2} \frac{\partial y(t_i, p)}{\partial p^\top} \right) \right] \nonumber \\
		&= \mathop{\mathbb{E}}_{Y|p} \left[ \sum_{k=1}^{n_t} \sum_{i=1}^{n_t} \left[ \left( \frac{Y(t_k) - y(t_k, p)}{\sigma_{t_k}^2} \frac{\partial y(t_k, p)}{\partial p} \right) \times \left( \frac{Y(t_i) - y(t_i, p)}{\sigma_{t_i}^2} \frac{\partial y(t_i, p)}{\partial p^\top} \right) \right] \right] \nonumber \\
		&= \sum_{k=1}^{n_t} \sum_{i=1}^{n_t} \mathop{\mathbb{E}}_{Y|p} \left[ \left( \frac{Y(t_k) - y(t_k, p)}{\sigma_{t_k}^2} \frac{\partial y(t_k, p)}{\partial p} \right) \times \left( \frac{Y(t_i) - y(t_i, p)}{\sigma_{t_i}^2} \frac{\partial y(t_i, p)}{\partial p^\top} \right) \right] \nonumber \\
		&= \sum_{k=1}^{n_t} \sum_{i=1}^{n_t} \frac{1}{\sigma_{t_k}^2} \frac{\partial y(t_k, p)}{\partial p} \frac{1}{\sigma_{t_i}^2} \frac{\partial y(t_i, p)}{\partial p^\top} \mathop{\mathbb{E}}_{Y|p} \left[ \left( Y(t_k) - y(t_k, p) \right) \times \left( Y(t_i) - y(t_i, p) \right) \right] 
	\end{aligned} 
	\end{equation} }

	Considering the mismatch between the dataset and the model output, the expectation $\mathop{\mathbb{E}}_{Y|p}$ can be redefined as $\mathop{\mathbb{E}}_{Y|p} \left[ \left( Y(t_i) - y(t_i, p) \right) \left( Y(t_k) - y(t_k, p) \right) \right] = \mathop{\mathbb{E}}_{Y|p} \left[ \epsilon_i \epsilon_k \right]$. For $i=k$, the expectation of $\epsilon_i$ is the variance of the noise $\sigma_{t_i}^2$. Analogously, the expectation becomes 0 if $i \neq k$ because the measurement noises are independent at different times.

	{\footnotesize
	\begin{equation}
		\mathop{\mathbb{E}}_{Y|p} \left[ \left( Y(t_i) - y(t_k, p) \right) \left( Y(t_i) - y(t_i, p) \right) \right] = \sigma_{t_i}^2 \delta_{ik}
	\end{equation} }

@@

	{\footnotesize
	\begin{equation}
		\text{FIM(p)} = \sum_{i=1}^{n_t} \left( \frac{1}{\sigma_{t_i}^2} \frac{\partial y(t_i, p)}{\partial p} \frac{\partial y(t_i, p)}{\partial p^\top} \right)
	\end{equation} }

@@

	{\footnotesize
	\begin{equation}
		\text{\text{FIM(p)}} = \frac{\partial y(t, p)}{\partial p} \begin{bmatrix}
			\frac{1}{\sigma_{t_1}^2} & 0 & 0\\
			0 & \ddots & 0 \\
			0 & 0 & \frac{1}{\sigma_{t_{n_t}}^2} 
		\end{bmatrix}\ \frac{\partial y(t, p)}{\partial p^\top} 
	\end{equation} }

	\subsubsection{Cramer-Rao inequality}

@@

	{\footnotesize
	\begin{equation}
		\Sigma = \mathop{\mathbb{E}}_{Y|p^*} \left[ \left( \hat{p}(Y) - p^* \right) \left( \hat{p}(Y) - p^* \right)^\top \right]
	\end{equation} }

	which quantifies how the estimates are spread around the true value $p^*$. One would like the estimates to be as concentrated as possible around this true value. An estimator $\hat{p}_1$ with covariance matrix $\Sigma_1$ is said to be more efficient than an estimator $\hat{p}_2$ with covariance matrix $\Sigma_2$ if $\Sigma_1 < \Sigma_2$, that is if $\Sigma_2 - \Sigma_1$ is positive-definite (i.e.if all the eigenvalues of $\Sigma_2-\Sigma_1$ arc strictly positive). Since estimators with high efficiency are desirable, a natural request is to make P as small as possible. The Cramer-Rao inequality provides a lower bound to what can be achieved.

@@

	\begin{itemize}
		\item the set of all data vectors $Y$ with $\pi_y(Y|p) > 0$ does not depend on $p$
		\item $\frac{\partial \pi_y(Y|p)}{\partial p_i}~\left(i=1,2,...,n_p\right)$ is absolutely integrable
		\item $\mathop{\mathbb{E}}_{Y|p} \left[ \frac{\partial \ln (\pi_y (Y|p))}{\partial p} \frac{\partial \ln (\pi_y (Y|p))}{\partial p^\top} \right]$ exists and is invertible
	\end{itemize}

@@

	{\footnotesize
	\begin{equation}
		\Sigma \geq \text{FIM(p)}^{-1}(p^*)
	\end{equation} }

@@

	\subsubsection{Optimal experimental design}

@@

	{\footnotesize
	\begin{equation}
		\Xi^* = \arg~\underset{\Xi \in \bar{\Xi}}{\text{opt}}~j\left(\Xi\right)
	\end{equation} }

@@

	{\footnotesize
	\begin{equation}
		j(\Xi) = \phi\left[ \text{FIM(p)}(p, \Xi) \right]
	\end{equation} }

@@

	{\footnotesize
	\begin{align}
		\phi_k(\text{FIM(p)}) &= \left[\frac{1}{n_p} \text{trace}\left( Q\text{FIM(p)}^{-1}Q^\top \right)^k \right]^{1/k} \quad &\text{if} \det F \neq 0 \nonumber \\
		\phi_k(\text{FIM(p)}) &= \infty \quad &\text{if} \det F = 0
	\end{align} }

@@

	{\footnotesize
	\begin{equation}
		j_L(\Xi) = \text{trace} \left[ Q^\top Q\text{FIM}^{-1}(p,\Xi) \right]
	\end{equation} }

	and choosing $Q = \textbf{I}_{n_p}$ then corresponds to the A-optimality cost function. An A-optimal experiment minimizes the sum of the squares of the lengths or the axes of asymptotic confidence ellipsoids. Choosing $Q$ diagonal, with $[Q]_{ii} = 1/p_i$ corresponds to C-optimality, which is connected with the relative precision of estimates. Taking Q to be a row vector leads to c-optimality. Taking $Q = \textbf{I}_{n_p}$ and $k = \infty$ corresponds to E-optimality; E-optimal design maximizes the smallest eigenvalues of the Fisher information matrix and thus minimizes the length of the largest axis of the asymptotic confidence ellipsoids. The most widely used optimality criterion has $k = 0$, $Q = \textbf{I}_{n_p}$ , requiring minimization of $\det \text{FIM}^{-1}(p, \Xi)$, or, equivalently, maximization of

	{\footnotesize
	\begin{equation}
		j_D(\Xi) = \det \text{FIM}(p, \Xi)
	\end{equation} }

@@

	\subsubsection{Problem formulation}

@@

	{\footnotesize
	\begin{equation}
		\begin{aligned} 
			&\Xi^* &= \arg \min_{ T^0, T^{in} \in \Xi} -j_D(\Xi,y)  \\
			&\text{subject to}
			& \dot{x} = f(x,t,\theta;\Xi) \\
			&& \dot{y} = g(x(t)) \\
			&& 40^\circ C \leq T^{in}(t) \leq 50^\circ C \\
			&& T^{0} = T^{in}(t=0) \\
		\end{aligned}
	\end{equation} }

	The non-linear optimization problem is solved by applying single-shooting to obtain the trajectory under assumption of piece-wise constant controls. The CasADi framework is used as the interface for the IPOPT solver. Each simulation problem is solved multiple time starting from different initial points and the solution with the best score is consider as the optimal solution.
\end{document}

### 2024-01-21 SFE_article_Design_of_Experiment/Sections/Porosity.tex
21.01

    {\footnotesize
        \begin{equation}
            \text{Volume excluding pores} = \dfrac{\text{mass}}{\text{true density}} = \dfrac{1\ \text{kg}}{1250\ \text{kg/m}^3} 0.0008\ \text{m}^3
        \end{equation}
    }

@@

    {\footnotesize
        \begin{equation}
            \phi = \dfrac{0.0016\ \text{m}^3 - 0.0008\ \text{m}^3}{0.0016\ \text{m}^3} = 0.5
        \end{equation}
    }

    Therefore, the porosity of the solid particles is 0.5 or 50\%. This means that half of the total volume of the solid particles is made up of pore space.

>>>>

    {\footnotesize
        \begin{equation}
            \text{Volume excluding pores} = \dfrac{\text{mass}}{\text{true density}} = \dfrac{1\ \text{kg}}{1256.58\ \text{kg/m}^3} 0.0007958\ \text{m}^3
        \end{equation}
    }

@@

    {\footnotesize
        \begin{equation}
            \phi = \dfrac{0.0016\ \text{m}^3 - 0.0007958\ \text{m}^3}{0.0016\ \text{m}^3} = 0.5026
        \end{equation}
    }

    Therefore, the porosity of the solid particles is 0.5026 or 50.26\%. This means that half of the total volume of the solid particles is made up of pore space.
